\documentclass[10pt]{article}
\usepackage{fontawesome}
\usepackage{setspace}
\onehalfspacing
\usepackage[colorlinks=true]{hyperref}
\usepackage[outputdir=build,newfloat]{minted}
\usepackage{mdframed}
\usepackage[margin=1.2in]{geometry}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{mathtools}
\theoremstyle{definition}
\newtheorem{question}{{\normalfont \faGears}~Question}
\newtheorem{compexercise}{{\normalfont \faLaptop}~Computer exercise}
\theoremstyle{remark}
\newtheorem*{protosolution}{Solution}
\usepackage{enumitem}
\usepackage{xcolor}
\setlist[enumerate]{font=\bfseries}
\input{../../macros}


\definecolor{lightgreen}{HTML}{f1faf8}
\newenvironment{solutionframe}
{%
    \begin{mdframed}[
        leftmargin=1cm,
        skipabove=.3cm,
        linecolor=blue,
        backgroundcolor=lightgreen,
        linewidth=0pt,
        innerleftmargin=.5em,
        innerrightmargin=.5em,
        innertopmargin=.3em,
        innerbottommargin=.6em,
    ]
}
{
    \end{mdframed}
}

\newenvironment{solution}
{\pushQED{\qed}\renewcommand{\qedsymbol}{$\triangle$}
\begin{solutionframe}\small \begin{protosolution}}
{\popQED\end{protosolution}\end{solutionframe}}

\begin{document}

\title{Numerical Analysis: Midterm \\
\small{(\textbf{40 marks})}}
\author{Urbain Vaes}
\maketitle

\begin{question}
    [Floating point arithmetic, \textbf{10 marks}]
    True or false?
    \begin{enumerate}
        \item
            Let $(\placeholder)_2$ denote binary representation.
            It holds that
            \(
                (0.1011)_2 + (0.0101)_2 = 1.
            \)

        \item
            Let $(\placeholder)_3$ denote base 3 representation.
            It holds that
            \(
                (1000)_3 \times (0.002)_3 = 2.
            \)

        \item
            A natural number with binary representation $(b_4 b_3 b_2 b_1 b_0)_2$ is even if and only if $b_0 = 0$.

        \item
            In Julia, \julia{Float64(.4) == Float32(.4)} evaluates to \julia{true}.

        \item
            Machine addition~$\madd$ is a commutative operation.
            More precisely, given any two double-precision floating point numbers $x \in \floating_{64}$ and $y \in \floating_{64}$,
            it holds that
            \(
                x \madd y = y \madd x.
            \)

        \item
            Let $\floating_{32}$ and $\floating_{64}$ denote respectively the sets of single and double precision floating point numbers.
            It holds that $\floating_{32} \subset \floating_{64}$.

        \item
            In Julia, \julia{eps(Float16)} returns the smallest strictly positive number that can be represented exactly in the \julia{Float16} format.

        \item
            Let $\floating_{64}$ be the set of double precision floating point numbers.
            For any $x \in \real$ such that~$x \in \floating_{64}$,
            it holds that $x + 1 \in \floating_{64}$.

        \item
            Let $x \in \real$ and $y \in \real$ be two numbers that are exactly representable in the \julia{Float64} format.
            Then $x \madd y = x + y$: machine addition is exact in this case.

        \item
            It holds that $(0.\overline{2200})_3 = (0.9)_{10}$.
    \end{enumerate}
\end{question}

\newpage
\begin{question}
    [Interpolation and approximation, \textbf{10 marks}]
    Are the following assertions true or false?
    Throughout this exercise, we use the notation
    $x^n_i = i/n$.
    The notation~$\poly(n)$ denotes the set of polynomials of degree less than or equal to~$n$.
    We proved in class that,
    for any function $u \colon \real \to \real$ and for all $n \in \nat_{>0}$,
    there exists a unique polynomial~$p_n \in \poly(n)$ such that
    \begin{equation}
        \label{eq:interpolation}
        \forall i \in \{0, \dotsc, n\}, \qquad
        p_n(x^n_i) = u (x^n_i).
    \end{equation}

    \begin{enumerate}
        \item
            If $u$ is not the zero function,
            then the degree of $p_n$ is exactly~$n$.

        \item
            If $u(x) = 2x + 1$,
            then $p_n = u$ for all $n \in \{1, 2, 3, \dotsc\}$.

        \item
            Fix $u(x) = 1 + \sin(57\pi x)$. Then $p_3(x) = 1$.

        \item
            Fix $u(x) = (2x - 1)^3$.
            Then $p_2(x) = 2x - 1$.

        \item
            Fix $n \in \nat_{>0}$ and suppose that $u\colon \real \to \real$ is a smooth function.
            There exists a constant~$K > 0$ independent of~$x$ such that
            \[
                \forall x \in \real,
                \qquad u(x) - p_n(x) = K \prod_{i=0}^{n} \left(x - x^n_i\right).
            \]

        \item
            It holds that
            \[
                \forall x \in [0, 1], \qquad
                \Bigl\lvert (x - x^n_0) \dotsc (x - x^n_n) \Bigr\rvert 
                \leq n! \left(\frac{1}{n}\right)^{n+1}.
            \]

        \item
            In Julia, assuming that $n$ and the function $u$ have already been defined,
            the following code enables to calculate the interpolating polynomial~$p_n$ of $u$:
            \begin{minted}{julia}
    using Polynomials
    # Assume `n=5` and `u` have already been defined
    x = LinRange(0, 1, n + 1)
    p = fit(x, u)
            \end{minted}

        \item
            Let $\Delta$ denote the finite difference operator:
            for a function $f \colon \real \to \real$,
            the function $\Delta f \colon \real \to \real$ is defined as
            \[
                \Delta f(x) = f(x + 1) - f(x).
            \]
            Then $f \in \poly(n)$ if and only if $\Delta^{n+1} f = 0$.
            Here $\Delta^{n+1}$ denotes the composition of $n+1$ applications of the operator~$\Delta$.

        \item
            In Julia, the following code enables to fit the data \julia{x} and \julia{y} by a straight line.
            \begin{minted}{julia}
    using Polynomials
    x = [1, 2, 3, 4]
    y = [16, 9, 4, 1]
    p = fit(x, y, 1)
            \end{minted}

        \item
            Using Chebyshev instead of equidistant points can improve on the quality of the interpolation.
\end{enumerate}
\end{question}

\newpage
\begin{question}
    [Numerical integration, \textbf{10 marks}]
    The Gauss--Legendre quadrature formula with~$n$ nodes is an approximate integration formula of the form
    \begin{equation}
        \label{gauss_legendre}
        I(u) := \int_{-1}^{1} u(x) \, \d x \approx \sum_{i=1}^{n} w_i \, u(x_i) =: \widehat I_n(u),
    \end{equation}
    which is exact when $u$ is a polynomial of degree less than or equal to $2n-1$.
    (Note that the nodes are here numbered starting from 1.)

    \begin{enumerate}
        \item
            \mymarks{5}
            Find the nodes and weights of the Gauss--Legendre rule with $n=3$ nodes,
            without using any formula (unless you prove it beforehand).

            \textbf{Hint:}
            Recall that a necessary and sufficient condition in order for~\eqref{gauss_legendre} to be satisfied for any polynomial~$p \in \poly(5)$ is that
            \[
                \int_{-1}^{1} x^d \, \d x= \sum_{i=1}^{n} w_i x_i^d,
                \qquad \text{ for all $d \in \{0, 1, 2, 3, 4, 5\}$}.
            \]
            Furthermore, given the symmetry of the problem,
            it is reasonable to look for a solution of the following form,
            which enables to reduce the number of unknowns.
            \[
                (x_1, x_2, x_3, w_1, w_2, w_3)
                = (-x, 0, x, w_1, w_2, w_1).
            \]

        \item \mymarks{5} Are the following assertions true of false :
            \begin{itemize}
                \item
                    The degree of precision of the composite trapezium rule is~$2$.

                \item
                    The composite Simpson rule can be used to integrate exactly a quadratic polynomial.

                \item
                    The degree of precision of the following rule is~$1$:
            \begin{minted}{julia}
    function my_integrate(f, a, b)
        x = LinRange(a, b, 100)
        h = x[2] - x[1]
        return h * sum(f, x[1:end-1])
    end
            \end{minted}

                \item
                    The degree of precision of the following integration rule is 2:
                    \[
                        \int_{-1}^{1} f(x) \, \d x \approx
                        2 f(0) + \frac{1}{3} f''(0).
                    \]

                \item
                    Suppose that $u \colon \real \to \real$ is a smooth function,
                    and let $\widehat I_n(u)$ denote an approximation of
                    the integral $I(u) := \int_{-1}^{1} u(x) \, \d x$
                    by the composite trapezium approximation with $n$ points.
                    Let
                    \[
                        \widehat J_n(u) = 2 \widehat I_{2n}(u) - \widehat I_n(u).
                    \]
                    It holds that
                    \[
                        \lim_{n \to \infty} n^2 \left\lvert I(u) - \widehat J_n(u) \right\rvert
                        = 0.
                    \]
            \end{itemize}

    \end{enumerate}
\end{question}

\begin{compexercise}
    [Interpolation, \textbf{10 marks}]
    Consider the following data:
    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c|}
            \hline
            \textbf{Time (hours)} & \textbf{Temperature (°C)} \\
            \hline
            6  & 10.5 \\
            \hline
            9  & 15.0 \\
            \hline
            12 & 20.2 \\
            \hline
            15 & 25.1 \\
            \hline
            18 & 22.8 \\
            \hline
            21 & 17.4 \\
            \hline
        \end{tabular}
        \caption{Recorded temperatures at different times of the day.}
    \end{table}

    We wish to approximate the temperature as a smooth function of time.
    To this end, calculate the interpolating polynomial~$p_{\rm int}$, 
    as well as the best quadratic polynomial approximation~$p_{\rm app}$ (in the sense that the sum of square errors is minimized).
    You may use the \julia{Polynomials} library.
    Plot on the same graph:
    \begin{itemize}
        \item The data points using \julia{scatter};
        \item The polynomial $p_{\rm int}$ interpolating the data points;
        \item The quadratic polynomial $p_{\rm app}$ that best approximates the data,
            in the sense of least squares.
    \end{itemize}
\end{compexercise}

\begin{compexercise}
    [Numerical integration, \textbf{10 marks}]
    Boole's integration rule reads
    \[
        \int_{-1}^{1} u(x) \, \d x
        \approx \frac{7}{45} u(-1) + \frac{32}{45} u\left(-\frac{1}{2}\right) + \frac{12}{45} u\left(0\right) + \frac{32}{45} u\left(\frac{1}{2}\right) + \frac{7}{45} u(1).
    \]
    \begin{itemize}
        \item
            Write a function \julia{comp_boole(u, a, b, N)},
            which returns an approximation of the integral
            \[
                I(u) = \int_{a}^{b} u(x) \, \d x
            \]
            obtained by partitioning the integration interval $[a, b]$ into $N$ cells,
            and applying Boole's rule within each cell.

        \item
            Take $u(x) = \cos(x)$, $a = -1$ and $b = 1$.
            Plot the evolution of the error for $N \in \{1, \dotsc, 200\}$.

        \item
            Estimate the order of convergence with respect to $N$, i.e.\ find~$\alpha$ such that
            \[
                \lvert \widehat I_{N} - I \rvert \propto C N^{-\alpha},
            \]
            where $I$ denotes the exact value of the integral
            and $\widehat I_{N}$ denotes its approximation by the composite Boole's rule.
            Use the method you prefer in order to find~$\alpha$.
            You can, for example, 
            use the function \julia{fit} from the \julia{Polynomials} package to find a linear approximation
            of the form
            \[
                \log \lvert \widehat I_{N} - I \rvert \approx \log (C) - \alpha \log(N).
            \]
    \end{itemize}
\end{compexercise}

\end{document}
